---
title: "Untitled"
output: html_document
date: "2025-11-17"
---

```{r}
library(tidyverse)
library(testthat)

options(repr.plot.width = 9, repr.plot.height = 6)
theme_set(theme_gray(base_size = 18))
```

## Question 1

*Use the cars dataframe to create a scatter plot with Weight_in_lbs on the x-axis and Horsepower on the y-axis. Color the points by the Origin of the cars.*

```{r}
cars <- read_csv('../data/cars.csv')
head(cars)
```

```{r}
ggplot(data = cars, aes(x = Weight_in_lbs, y = Horsepower, color = Origin )) + 
  geom_point() +
  labs(x = 'Weight in Lbs', y = 'Horsepower') +
  theme_minimal()
```

## 

*Let’s use a line plot to visualize how the price of a few common stocks has changed over time. Create a line plot with the stocks’ date on the x-axis and their price on the y-axis. You should colour the lines by the stock symbol so that you can compare the stock development for the different companies.*

```{r}
stocks <- read_csv('../data/stocks.csv')
glimpse(stocks)

head(stocks)
```

```{r}
ggplot(data = stocks, aes(x = date, y = price, color = symbol)) +
  geom_line()+
  labs(title = 'Stock Prices', x= 'Date', y = 'Prices')+
  theme_minimal()
```

I have uploaded the 2018 Gapminder data at this URL and assign it a suitablevariable name. Use read_csv from readr to load the data directly from the URL.

*Now let’s create a similar bubble chart to what you saw in the video: Filter the dataframe to only keep observations from a single year, 1962. You can create a new data frame variable or (preferably) perform the filtering directly and pipe it to ggplot. Dates can be matched as strings when filtering. Use a point geom to recreate the appearance of the plot in the video. Encode the proper variables so that children per woman is on the x-axis, life expectancy on the y-axis, and so that the circles’ color corresponds to their region, and the size reflects the population.*

*Don't worry about getting axis labels and sizes to be exactly like in the video, we will return to this code later in the lab to customize it.*

```{r}
url <- 'https://raw.githubusercontent.com/UofTCoders/workshops-dc-py/master/data/processed/world-data-gapminder.csv'

gap <- read_csv(url) |> 
  filter(year == 1962) |> 
  ggplot(aes(x = children_per_woman, y = life_expectancy, color = region, size = population)) +
  geom_point()+
  labs(x= 'Children per Woman', y= 'Life Expectancy')+
  theme_minimal()

gap
```
